---
title: "Reproducing data in squidpy package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing data in squidpy package}
  %\VingetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction

This vignette will demonstrate to the user how to reproduce the MERFISH data
that was used in the squidpy manuscript. The authors of squidpy documented their
work via a Jupyter notebook that we then translated into R code. The steps below
use purely R functions / objects to obtain the MERFISH data object that is
present in the squidpy package.

<!-- TODO: Add references to manuscript and github for squidpy -->

# Data preperation

First, users will need to download the raw data which we found
[here](https://datadryad.org/stash/dataset/doi:10.5061/dryad.8t8s248). We
download the data and then do some data manipulation. The steps are outlined
below.

## Download data

```{r download_data, eval = FALSE}
library(rdryad)
dat_path <- dryad_download("10.5061/dryad.8t8s248")
counts <- read.csv(dat_path[[1]])
``` 

## Subset data

We subset the full data set into metadata and counts, then retrieve the genes.

```{r subset}
metadata <- counts[,1:9]
counts_g <- counts[,10:170]

genes <- colnames(counts_g)
```

## Creating the csr matrix

First we replace all NA with 0 then create a csr matrix, dropping all 0
elements.

```{r csr_matrix}
counts_g <- replace(counts_g, is.na(counts_g), 0)

library(Matrix)
mat <- data.matrix(counts_g)
counts_g_sparse <- as(drop0(mat), "RsparseMatrix")
```

## Creating a SingleCellExperiment object

In the Jupyter notebook the authors choose to create an AnnData object. We at
Bioconductor prefer to represent the data as a SummarizedExperiment or
SingleCellExperiment object. The primary difference being the composition of the
data. We prefer to have the data represented with the genes as rows and the
samples as columns. Then we add the spatial data as a reduced dimension.

```{r sce}
library(SingleCellExperiment)
sce <- SingleCellExperiment(
    assays = list(counts = t(counts_g_sparse)),
    colData = metadata
)

spatial <- colData(sce)[,c("Centroid_X", "Centroid_Y")]
reducedDim(sce, "spatial") <- spatial
```

# Notebook continued

From here we try to follow the Jupyter notebook exactly using our sce object
instead of their anndata object. Most is just data transformation. The end
result should be the same as what is included in the squpidpy package.

```{r notebook}
sce$Bregma = sce$Bregma * 100

unique(sce$Animal_ID)

sce_filt <- sce[,sce$Animal_ID == 1]
sort(unique(sce_filt$Bregma))

sce_list <- c()
spatial_list <- c()
for (i in sort(unique(sce_filt$Bregma))) {
    s_filt <- sce_filt[sce_filt$Bregma == i]
    sce_list <- c(sce_list, s_filt)
    spatial_filt <- reducedDim(s_filt)
}
```
